Miniproject - Combinatorics

### 2021-12-03 Combinatorics/submission.tex
Added some computations for the size of L_i of multi-subsets of {0, ..., k}^d

    Assume not, then 
    \begin{equation*}
        r(a_i) - r(a_{i - 1}) = \sum_j a_{ij} - a_{(i - 1)j} \geq 2
    \end{equation*}
    For all $j$ we have $a_{ij} \geq a_{(i - 1)j}$ and so there is $s \leq d$ with $a_{is} - a_{(i - 1)s} \geq 2$ or there are different $r, s \leq d$ with $a_{ir} - a_{(i - 1)r}, a_{is} - a_{(i - 1)s} \geq 1$.
    In both cases, find
    \begin{equation*}
        \tilde{a} \in \mathcal{P}_{k, d} \ \text{defined by} \ \tilde{a}_j = \begin{cases}
            a_{(i - 1)j} & \text{if $j \neq s$} \\
            a_{(i - 1)j} + 1 & \text{otherwise}
        \end{cases}
    \end{equation*}
    with $a_{i - 1} < \tilde{a} < a_i$. However, this gives a longer chain
    \begin{equation*}
        a_1 < ... < a_{i - 1} < \tilde{a} < a_i < ... < a_t
    \end{equation*}
    which contradicts the assumed maximality. The claim follows.
\end{proof}
\end{document}

>>>>

    Assume not, then 
    \begin{equation*}
        r(a_i) - r(a_{i - 1}) = \sum_j a_{ij} - a_{(i - 1)j} \geq 2
    \end{equation*}
    For all $j$ we have $a_{ij} \geq a_{(i - 1)j}$ and so there is $s \leq d$ with $a_{is} - a_{(i - 1)s} \geq 2$ or there are different $r, s \leq d$ with $a_{ir} - a_{(i - 1)r}, a_{is} - a_{(i - 1)s} \geq 1$.
    In both cases, find
    \begin{equation*}
        \tilde{a} \in \mathcal{P}_{k, d} \ \text{defined by} \ \tilde{a}_j = \begin{cases}
            a_{(i - 1)j} & \text{if $j \neq s$} \\
            a_{(i - 1)j} + 1 & \text{otherwise}
        \end{cases}
    \end{equation*}
    with $a_{i - 1} < \tilde{a} < a_i$. However, this gives a longer chain
    \begin{equation*}
        a_1 < ... < a_{i - 1} < \tilde{a} < a_i < ... < a_t
    \end{equation*}
    which contradicts the assumed maximality. The claim follows.
\end{proof}
\begin{definition}
    Let $P$ be a graded poset with layers $L_0, ..., L_n$.
    A symmetric chain in $P$ is a chain $C \subseteq P$ such that there is $i \leq n/2$ with
    \begin{equation*}
        \forall 0 \leq j \leq n: \ C \cap L_j \neq \emptyset \ \Leftrightarrow \ i \leq j \leq n - i
    \end{equation*}
    A decomposition $\mathcal{C}$ of $P$ into chains is called symmetric chain decomposition, if all $C \in \mathcal{C}$ are symmetric.
\end{definition}
Note that this definition is compatible with the definition of a symmetric chain for the classical graded poset $\powerset(n)$.
\begin{proposition}
    Let $k, d \geq 1$. Then $\mathcal{P}_{k, d}$ has a symmetric chain decomposition.
\end{proposition}
\begin{proof}

\end{proof}
\begin{proposition}
    Let $M(k, d, n) := |\{ x \in \mathcal{P}_{k, d} \ | \ r(x) = n\}|$. Then for $k, d \geq 1$ we have the recurrence relation
    \begin{equation*}
        M(k, 1, n) = \begin{cases}
            1 & \text{if $0 \leq n \leq k$} \\
            0 & \text{otherwise}
        \end{cases}
        \quad \text{and} \quad
        M(k, d, n) = \sum_{i = 0}^k M(k, d - 1, n - i)
    \end{equation*}
\end{proposition}
\begin{corollary}
    Let $k, d \geq 1$. Then
    \begin{align*}
        M(k, 2, n) := \begin{cases}
            0 & \text{if $n > 2k$} \\
            k + 1 & \text{if $k < n \leq 2k$} \\
            n + 1 & \text{if $0 \leq n \leq k$} \\
            0 & \text{if $n < 0$}
        \end{cases}
    \end{align*}
    and
    \begin{align*}
        M(k, 3, n) := \begin{cases}
            0 & \text{if $n > 3k$} \\
            (k + 1)^2 & \text{if $2k < n \leq 3k$} \\
            (k + 1)(2n + 2 - k) / 2 & \text{if $k < n \leq 2k$} \\
            {n + 1 \choose 2} & \text{if $0 \leq n \leq k$} \\
            0 & \text{if $n < 0$}
        \end{cases}
    \end{align*}
\end{corollary}
\begin{proof}
    Just use the recurrence relation twice.
    The only nontrivial computation is if $k < n \leq 2k$ and $d = 3$; We find
    \begin{align*}
        &\sum_{i = 0}^k M(k, 2, n) = \sum_{i = 0}^n M(k, 2, n) = \sum_{i = 0}^{n - 2k - 1} 0 + \sum_{i = n - 2k}^{n - k - 1} (k + 1) + \sum_{i = n - k}^n (i + 1) \\
        =& (k + 1)(n - k - 1 - (n - 2k) + 1) + \frac {(n + 1)(n + 2)} 2 - \frac {(n - k)(n - k + 1)} 2 \\
        =& k(k + 1) + kn + n + 1 - \frac {k^2} 2 + \frac k 2 = \frac {(k + 1)(2n + 2 - k)} 2
    \end{align*}
\end{proof}
\end{document}
